---
title: "SPS data"
output: html_notebook
---

SPS

```{r libraries, include=FALSE}
source("libraries.R")
```
Louisa Boren STEM K-8	West Seattle HS + Denny Int. MS
Pathfinder K-8	West Seattle HS + Denny Int. MS
Orca K-8	Mercer Int MS + Hawthorne Elementary + Kimball Elementary
South Shore K8	Aki Kurose MS - Orca geozone
John Stanford International School	Hamilton Int. MS
TOPS K-8	Elementary attendance areas of: Montlake, McGilvra, Stevens, Lowell(but not past I5), Madrona, Gatzert, Leschi, Thurgood Marshall, John Muir and Hawthorn
Thornton Creek Elementary	Eckstein MS + Olympic View Elementary
Hazel Wolf K-8	Jane Addams MS + Eckstein MS + Olympic View Elementary
Cedar Park	Jane Addams MS
Licton Springs K-8	Jane Adams MS + Eckstein MS + Robert Eagle Staff MS
Salmon Bay K-8	Hamilton Int. MS + Whitman MS + Robert Eaglestaff MS
McDonald International School	Hamilton Int. MS
Queen Anne Elementary	McClure Middle School

```{r}
attendance_areas <- st_read("~/SPS/SPS_AttendanceAreasAndSchools_Shapefiles_2020_2021/ES/sps_attendance_area_ES_2020_2021.shp")
school.sites <- st_read("~/SPS/SPS_AttendanceAreasAndSchools_Shapefiles_2020_2021/sites/Sites_2020.shp")


option.sites <- school.sites %>% 
  filter(property_l=="Orca K-8" | property_l=="Boren STEM K-8" | property_l=="Pathfinder K-8" | property_l=="South Shore K-8" | 
           property_l=="John Stanford Int'l" |  property_l=="TOPS K-8" |  property_l=="Leschi" |  property_l=="Thornton Creek" | property_l=="Hazel Wolf K-8" | 
         property_l=="Cedar Park" | property_l=="Licton Springs K-8" | property_l=="Salmon Bay K-8" | property_l=="McDonald Int'l" | property_l=="Queen Anne")

#Louisa Boren STEM K-8	Madison MS + Denny Int. MS

trans.boren <- attendance_areas %>% 
  filter(MS_ZONE=="Denny Int'l" | MS_ZONE=="Madison") %>% 
  st_combine() %>% 
  st_union(by_feature = TRUE)

  
#Pathfinder K-8	Madison MS + Denny Int. MS
trans.pathfinder <- attendance_areas %>% 
  filter(MS_ZONE=="Denny Int'l" | MS_ZONE=="Madison") %>% 
  st_combine() %>% 
   st_union(by_feature = TRUE)


#OrcaK-8 attendance area Mercer Int MS + Hawthorne Elementary + Kimball Elementary
trans.orca <- attendance_areas %>% 
  filter(MS_ZONE=="Mercer Int'l" | ES_ZONE=="Kimball" | ES_ZONE=="Hawthorne") %>% 
     st_combine() %>% 
   st_union(by_feature = TRUE)


# South Shore K8	Aki Kurose MS - Hawthorne elementary
trans.southshore <- attendance_areas %>% 
  filter(MS_ZONE=="Aki Kurose" | ES_ZONE=="Hawthorne") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# John Stanford International School	Hamilton Int. MS
trans.jsis <- attendance_areas %>% 
  filter(MS_ZONE=="Hamilton Int'l") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# TOPS K-8	Elementary attendance areas of: Montlake, McGilvra, Stevens, Lowell(but not past I5), Madrona, Gatzert, Leschi, Thurgood Marshall, John Muir and Hawthorn
trans.tops <- attendance_areas %>% 
  filter(ES_ZONE=="Montlake" | ES_ZONE=="McGilvra" | ES_ZONE=="Stevens" |
           ES_ZONE=="Lowell" | ES_ZONE=="Madrona" | ES_ZONE=="Gatzert" | ES_ZONE=="Leschi" |
ES_ZONE== "Thurgood Marshall" | ES_ZONE =="John Muir"| ES_ZONE=="Hawthorne") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# Thornton Creek Elementary	Eckstein MS + Olympic View Elementary
trans.thorton <- attendance_areas %>% 
  filter(MS_ZONE=="Eckstein" | ES_ZONE=="Olympic View") %>% 
   st_combine() %>% 
   st_union(by_feature = TRUE)

# Hazel Wolf K-8	Jane Addams MS + Eckstein MS + Olympic View Elementary
trans.hazel <- attendance_areas %>% 
  filter(MS_ZONE=="Jane Addams" | MS_ZONE=="Eckstein" | ES_ZONE=="Olympic View") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# Cedar Park	Jane Addams MS
trans.cedarpark <- attendance_areas %>% 
  filter(MS_ZONE=="Jane Addams") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# Licton Springs K-8	Jane Adams MS + Eckstein MS + Robert Eagle Staff MS
trans.licton <- attendance_areas %>% 
  filter(MS_ZONE=="Jane Addams" | MS_ZONE=="Eckstein" | MS_ZONE=="Eagle Staff") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# Salmon Bay K-8	Hamilton Int. MS + Whitman MS + Robert Eaglestaff MS
trans.salmonbay <- attendance_areas %>% 
  filter(MS_ZONE=="Hamilton Int'l"| MS_ZONE=="Whitman" | MS_ZONE=="Eagle Staff") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# McDonald International School	Hamilton Int. MS
trans.mcdonald <- attendance_areas %>% 
  filter(MS_ZONE=="Hamilton Int'l") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

# Queen Anne Elementary	McClure Middle School
trans.queenane <- attendance_areas %>% 
  filter(MS_ZONE=="McClure") %>% 
    st_combine() %>% 
   st_union(by_feature = TRUE)

seattle.crs <- st_crs(trans.boren)


origin.language <- st_read("ACS%3A_Nativity_By_Language_Spoken_At_Home_By_Ability_To_Speak_English_For_The_Population_5_Years_And_Over___acs_b16005_nativitybylanguage.shp")


school.enrollment <- st_read("ACS%3A_School_Enrollment_By_Level_Of_School_For_The_Population_3_Years_And_Over___acs_b14001_schoolenrollment.shp")
#Attribute Label: E14001006
#Attribute Definition: E1216 - Enrolled in school:Enrolled in grade 5 to grade 8
#Attribute Label: E14001005
#Attribute Definition: E1216 - Enrolled in school:Enrolled in grade 1 to grade 4
#Attribute Label: E14001004
#Attribute Definition: E1216 - Enrolled in school:Enrolled in kindergarten

KCdemographics <- st_read("Consolidated_Demographics_Index_for_King_County_Census_Tracts___demographic_index_area.shp")

kc.demographics.nad <- st_transform(KCdemographics, seattle.crs)
school.enrollment.nad <- st_transform(school.enrollment, seattle.crs)
origin.language.nad <- st_transform(origin.language, seattle.crs)


get_demoinfo <- function(thisnoschool, school.transareas, school.names, city.region, kc.demographics.nad, school.enrollment.nad, origin.language.nad){

  print(thisnoschool)
  thistransportationarea <- school.transareas[[thisnoschool]]
  thisschool <- school.names[thisnoschool]
  thisregion <- city.region[thisnoschool]
  
  print(paste(thistransportationarea, thisschool, thisregion))
    
  this.demo <- st_intersection(kc.demographics.nad,thistransportationarea)
  
 this.enroll <- st_intersection(school.enrollment.nad,thistransportationarea)
 
 this.lang <- st_intersection(origin.language.nad,thistransportationarea)
 
 this.enroll.data <- this.enroll %>% 
  st_drop_geometry() %>% 
  as_tibble() %>% 
  mutate(enrolled_K_8 = E14001006 + E14001005 + E14001004) %>% 
  mutate(school = thisschool) %>% 
  group_by(school) %>% 
  summarise(tot_enrolled_K_8 = sum(enrolled_K_8)) %>% 
   mutate(region=thisregion)
   

 this.lang.data <- this.lang %>% 
  st_drop_geometry() %>% 
  as_tibble() %>% 
   mutate(language = E16005024 + E16005019) %>% 
    mutate(school = thisschool) %>% 
  group_by(school) %>%
   summarise(tot_lang = sum(language))
   
 
  area.trans <- st_area(thistransportationarea) %>% 
    as.numeric %>% 
    multiply_by(3.5870064279155E-8) 

this.demo.data <- this.demo %>% 
  st_drop_geometry() %>% 
  as_tibble() %>% 
  mutate(school = thisschool) %>% 
  group_by(school) %>% 
  summarise(Tot_population = sum(Population), Per_ESL = (sum(NumberofES)/sum(Population))*100, mean_Median_house_income = mean(MedianHous), Per_POC = (sum(NumberofPO)/sum(Population))*100) %>% mutate(transportation_area_sqmiles = area.trans) %>% 
  left_join(this.enroll.data, by="school") %>% 
  left_join(this.lang.data, by="school") %>% 
  mutate(Per_enrolled_K_8 = (tot_enrolled_K_8/Tot_population)*100) %>% 
  mutate(Per_other_language = (tot_lang/Tot_population)*100)


return(this.demo.data)
  
}

school.transareas <- list(trans.boren, trans.cedarpark, trans.hazel, trans.jsis, trans.licton, trans.mcdonald, trans.orca, trans.pathfinder, trans.queenane, trans.salmonbay, trans.southshore, trans.thorton, trans.tops)

school.names <- c("Louisa Boren STEM K-8","Cedar Park", "Hazel Wolf K-8", "John Stanford International School", "Licton Springs K-8", "McDonald International School", 
                  "Orca K-8", "Pathfinder K-8", "Queen Anne Elementary", "Salmon Bay K-8", "South Shore K8", "Thornton Creek Elementary", "TOPS K-8")

city.region <- c("West Seattle","North Seattle","North Seattle","North Seattle","North Seattle","North Seattle",
                 "Central Seattle", "West Seattle","Central Seattle","North Seattle","South Seattle","North Seattle","Central Seattle")

no.schools <- 1:length(school.names)

school.trans.demo <- lapply(no.schools, get_demoinfo, school.transareas, school.names, city.region, kc.demographics.nad, school.enrollment.nad,origin.language.nad)

school.trans.demo %>% 
  bind_rows %>% 
  arrange(region,transportation_area_sqmiles,school,Per_POC) %>% 
  write_csv("school_census.csv")


option.demo <- read_csv("option_school_demographics.csv") %>% 
  dplyr::rename(school=School)
  
school.sum.data <- school.trans.demo %>% 
  bind_rows %>% 
  arrange(region,transportation_area_sqmiles,school,Per_POC) %>% 
  left_join(option.demo, by="school")

write_csv(school.sum.data, "school_option_data.csv")

trans.income <- ggplot(data=school.sum.data) +
  geom_point(aes(transportation_area_sqmiles,mean_Median_house_income, color = region, fill = region)) +
    xlab("Transportation area Sq Miles") +
    ylab("Mean median household income") + 
  theme_economist() 

  ggsave("transportation_income.png",trans.income,device="png",dpi=300)
    
trans.esl <- ggplot(data=school.sum.data) +
  geom_point(aes(transportation_area_sqmiles,Per_ESL, color = region, fill = region)) +
    xlab("Transportation area Sq Miles") +
    ylab("Percent population speak English as a Second language") + 
  theme_economist() 

  ggsave("transportation_esl.png",trans.esl,device="png",dpi=300)
  
      
trans.pop <- ggplot(data=school.sum.data) +
  geom_point(aes(transportation_area_sqmiles, `FRL students elegible for busing`, color = region, fill = region)) +
    xlab("Transportation area Sq Miles") +
    ylab("% Free and Reduced Lunch School eligible busing") + 
  theme_economist() 

  ggsave("transportation_pop.png",trans.pop,device="png",dpi=300)


trans.ell <- ggplot(data=school.sum.data) +
  geom_point(aes(`%student pop. eleigle for busing`, `%FRL`, color = region, fill = region)) +
    xlab("% Student population elegible for busing") +
    ylab("% School population elegible for Free and Reduced Lunch") + 
  theme_economist() 

  ggsave("transportation_ell.png",trans.ell,device="png",dpi=300)
  
  
trans.bipoc <- ggplot(data=school.sum.data) +
  geom_point(aes(`%student pop. eleigle for busing`, `%BIPOC`, color = region, fill = region)) +
    xlab("% Student population elegible for busing") +
    ylab("% BIPOC Student population") + 
  theme_economist() 

  ggsave("transportation_bipoc.png",trans.bipoc,device="png",dpi=300)


  trans.area.bipoc <- ggplot(data=school.sum.data) +
  geom_point(aes(transportation_area_sqmiles, `%BIPOC`, color = region, fill = region)) +
    xlab("Transportation area Sq Miles") +
    ylab("% BIPOC Student population") + 
  theme_economist() 

  ggsave("transportation_area_bipoc.png",trans.bipoc,device="png",dpi=300)


```


```{r}

#Attribute:
#Attribute Label: E1601126, Attribute Definition: E1317 -Percent, SPEAK A LANGUAGE OTHER THAN ENGLISH - Spanish - 5 to 17 years old
#Attribute Label: M1601126, Attribute Definition: M131717 -Percent, SPEAK A LANGUAGE OTHER THAN ENGLISH - Spanish - 5 to 17 years old. Attribute Definition Source: US Census Margin of Error
#Attribute Label: E1601119, Attribute Definition: E1317 -Total, SPEAK A LANGUAGE OTHER THAN ENGLISH - Spanish
#Attribute Label: E1601113, Attribute Definition: E1317 -Total, Speak a language other than English
#Attribute Label: E1601101, Attribute Definition: M131717 -Total, Population 5 years and over

attendance_areas <- st_read("~/SPS/SPS_AttendanceAreasAndSchools_Shapefiles_2020_2021/ES/sps_attendance_area_ES_2020_2021.shp")
school.sites <- st_read("~/SPS/SPS_AttendanceAreasAndSchools_Shapefiles_2020_2021/sites/Sites_2020.shp")

language_home <- st_read("ACS%3A_Language_Spoken_At_Home___acs_s1601_languagespokeathome.shp")


ggplot() +
   geom_sf(aes(fill=(E1601119/E1601101)*100),data = language_home)

```

